\Huge Handwritten digit recognition

### 2017-02-09 doc/Paper.tex
update paper, add ss to readme

\usepackage[automark]{scrpage2} % heading
  \pagestyle{scrheadings}
  \ihead[]{Handwritten character recognition in MPI} % left header
  \ohead[]{\today} % right header
  \cfoot[]{\pagemark} % footer
  \setheadsepline[122mm]{0.3mm}

@@

\newpage
\section{Introduction}
Proiectul are ca scop evidentierea rolului paralelismului in programare. Avand vedere cantitatea mare de date cu care se confrunta aplicatiile din ziua de astazi, este nevoie ca acestea sa ruleze anumite sarcini concomitent. Pentru a sublinia imbunatatirile aduse de paralelism, s-a implementat in MPI antrenarea unei retele neuronale. In continuare se vor detalia, datele folosite in model, reteaua construita, metodele de paralelism, rezultatele, dar si interfata.

\section{Problem statement}
\label{sec:problem}
Pentru o retea neurala care foloseste un algoritm de antrenare bazat pe metoda gradientului descendent si dispune de o cantitate mare de date de antrenare timpul ca reteaua sa ajunga la rezultate satisfacatoare poate dura chiar si zile pe o implementare secventiala
Ne propunem sa elaboram o implementare paralela a metodei gradientului descendent. Vom folosi pentru validarea implementarii o problema celebra in computer vision, cea a detectarii cifrelor scrise de mana.

For a neural network witch use for training the gradient descendent algorithm and dispose a large amount of training data, the time to reach convergence it can take days in a sequentially implementation.
We purpose to elaborate a parallel implementation of gradient descendent, and for validate that we use a common problem in computer vision: given a picture of a handwritten digit, label it accordingly 0 - 9.

@@

We use a feed-forward neural network with an one fully connected layer, which predict the digit that has been in input image.
For an image $x$, initialy $28 \times 28$, we flatten this array into a vector of $784 = 28 \times 28$.

\begin{align}
y &= softmax(Wx + b) \\
label &= argmax(y)
\end{align}

In $(1)$, $W$ is the weights matrix, $10 \times 784$ -- the number of classes(one for each digit) and the dimension of an image ($784$). $b$ is the bias vector, $1 \times 10$, corresponding to the free element.
\\

The activation function output a vector of probabilities:
$$softmax(x_i)=\frac{exp(x_i)}{\sum_j exp(x_j)}$$
Where $exp(x)$ represents $e^x$, Euler's constant. It \textit{normalizes} the vector, causing the greatest elements to have logarithmic relative value.
\\

$(2)$ With $y$ vector of probabilities, $argmax$ take the index of the maximum value and results the prediction label.

@@

\TODO{more about evaluation?}

\noindent The following chart illustrates model evolution, both on the test and on the training set, after each pass through the dataset.

\begin{center}
\begin{tikzpicture}
\begin{axis} [
  title={Training performance},
  xlabel={Epoch},
  ylabel={Accuracy (\%)},
  xmin=0, xmax=100,
  ymin=0, ymax=100,
  xtick={0,25,50,75,100},
  ytick={0,25,50,75,100},
  legend pos=south east,
  axis line style={black!65, line width=.1pt}, % thinner frame border
  legend style={draw=black!50, line width=.1pt}, % thinner frame border
  every x tick/.style={white}, % hide x ticks (make them white)
  every y tick/.style={white}, % hide y ticks (make them white)
  every axis plot/.append style={thick},
]

\addplot[]
  coordinates {
  (0,25)(25,65)(50,90)(100,95)
  };
\addplot[dash pattern=on 6pt off 4pt]
  coordinates {
  (0,25)(25,50)(50,80)(100,90)
  };

@@

\TODO{comment on chart}

>>>>

% \usepackage[automark]{scrpage2} % heading
%   \pagestyle{scrheadings}
%   \ihead[]{Handwritten character recognition in MPI} % left header
%   \ohead[]{\today} % right header
%   \cfoot[]{\pagemark} % footer
%   \setheadsepline[122mm]{0.3mm}

@@

\newpage
\section{Introduction}
The project aims to highlight the role of parallelism in programming, especially in Machine Learning tasks. Considering the vast amount of data today's applications have to deal with, it is paramount to run tasks concomitantly. To underline  improvements bought by parallelism, we implemented  a neural network training module in MPI.

We will present the details, data used, model architecture, parallelization strategy and an interactive interface.

\section{Problem statement}
\label{sec:problem}
Given a picture of a handwritten digit, label it accordingly 0 - 9.
This is a classic problem in Computer Vision on which we apply a neural network trained using gradient descent. We seek to reach satisfactory model performance and fast training time, using parallelism.

@@

We use a feed-forward neural network with an one fully connected layer:
\begin{align*}
y &= f(Wx + b) \\
label &= argmax(y)
\end{align*}

Where $W$ is the weights matrix, $10 \times 784$ -- the number of classes(one for each digit) and the dimension of an image ($784$). $b$ is the bias vector, $1 \times 10$, corresponding to the free element. $y$ is a vector of probabilities, from which argmax extracts the predicted labe
\\

The activation function output a vector of probabilities:
$$f(x_i)=\frac{exp(x_i)}{\sum_j exp(x_j)}$$
Where $exp(x)$ represents $e^x$, Euler's constant. It \textit{normalizes} the vector, causing the greatest elements to have logarithmic relative value.
\\

@@

\noindent The following chart illustrates model evolution, after each pass through the dataset, on the entire MNIST training (60 000) and test (10 000) dataset.

\begin{center}
\begin{tikzpicture}
\begin{axis} [
  title={Training performance},
  xlabel={Epoch},
  ylabel={Accuracy (\%)},
  xmin=0, xmax=250,
  ymin=0, ymax=100,
  enlarge x limits=0.05,
  xtick={0,50,100,150,200, 250},
  ytick={0,25,50,75,100},
  legend pos=south east,
  axis line style={black!65, line width=.1pt}, % thinner frame border
  legend style={draw=black!50, line width=.1pt}, % thinner frame border
  every x tick/.style={white}, % hide x ticks (make them white)
  every y tick/.style={white}, % hide y ticks (make them white)
  every axis plot/.append style={thick},
]

\addplot[dash pattern=on 6pt off 4pt]
  coordinates {
 (0, 25.542341)    (1, 47.745369)    (2, 66.106474)    (3, 79.268610)    (4, 81.409645)    (5, 82.697285)    (6, 83.599091)    (7, 84.318434)    (8, 84.974273)    (9, 85.386156)    (10, 85.663764)    (11, 86.105388)    (12, 86.463159)    (13, 86.705436)    (14, 86.973765)    (15, 87.296094)    (16, 87.391127)    (17, 87.672536)    (18, 87.834651)    (19, 87.989275)    (20, 88.216571)    (21, 88.300424)    (22, 88.423408)    (23, 88.635834)    (24, 88.689947)    (25, 88.783079)    (26, 88.917243)    (27, 89.086849)    (28, 89.001096)    (29, 89.200441)    (30, 89.248964)    (31, 89.334605)    (32, 89.433439)    (33, 89.565703)    (34, 89.670015)    (35, 89.733408)    (36, 89.835932)    (37, 89.940244)    (38, 89.940244)    (39, 90.016606)    (40, 90.037178)    (41, 90.124720)    (42, 90.079998)    (43, 90.175031)    (44, 90.219753)    (45, 90.365097)    (46, 90.448950)    (47, 90.372588)    (48, 90.560753)    (49, 90.534703)    (50, 90.560753)    (51, 90.687426)    (52, 90.603574)    (53, 90.763900)    (54, 90.821590)    (55, 90.810410)    (56, 90.843951)    (57, 90.920425)    (58, 90.901753)    (59, 90.914834)    (60, 90.948375)    (61, 91.037818)    (62, 91.007967)    (63, 91.136541)    (64, 91.125360)    (65, 91.220393)    (66, 91.147721)    (67, 91.233474)    (68, 91.328507)    (69, 91.334097)    (70, 91.384409)    (71, 91.255835)    (72, 91.330296)    (73, 91.438410)    (74, 91.348967)    (75, 91.430919)    (76, 91.522262)    (77, 91.542722)    (78, 91.546524)    (79, 91.531542)    (80, 91.596835)    (81, 91.555803)    (82, 91.624786)    (83, 91.643345)    (84, 91.783100)    (85, 91.768230)    (86, 91.712328)    (87, 91.727198)    (88, 91.818541)    (89, 91.840902)    (90, 91.755149)    (91, 91.900493)    (92, 91.846492)    (93, 91.870754)    (94, 91.909885)    (95, 91.971377)    (96, 91.909885)    (97, 92.034657)    (98, 91.904295)    (99, 91.973165)    (100, 91.982557)    (101, 92.010508)    (102, 92.006706)    (103, 92.012297)    (104, 92.083180)    (105, 92.045838)    (106, 92.131591)    (107, 92.213543)    (108, 92.072000)    (109, 92.189393)    (110, 92.146461)    (111, 92.254575)    (112, 92.245295)    (113, 92.185592)    (114, 92.293706)    (115, 92.273246)    (116, 92.258264)    (117, 92.381248)    (118, 92.278836)    (119, 92.474492)    (120, 92.358887)    (121, 92.375658)    (122, 92.476281)    (123, 92.448330)    (124, 92.351508)    (125, 92.511722)    (126, 92.425969)    (127, 92.388739)    (128, 92.360788)    (129, 92.450231)    (130, 92.502443)    (131, 92.565723)    (132, 92.543363)    (133, 92.621625)    (134, 92.647787)    (135, 92.532182)    (136, 92.560133)    (137, 92.614246)    (138, 92.629116)    (139, 92.578804)    (140, 92.696198)    (141, 92.703689)    (142, 92.586295)    (143, 92.714869)    (144, 92.727838)    (145, 92.657067)    (146, 92.737230)    (147, 92.714869)    (148, 92.660756)    (149, 92.711068)    (150, 92.726050)    (151, 92.735329)    (152, 92.843443)    (153, 92.789330)    (154, 92.828461)    (155, 92.742820)    (156, 92.766970)    (157, 92.876984)    (158, 92.875083)    (159, 92.722248)    (160, 92.794920)    (161, 92.871394)    (162, 92.826673)    (163, 92.878773)    (164, 92.938476)    (165, 92.882574)    (166, 92.958936)    (167, 92.882574)    (168, 92.942165)    (169, 92.895543)    (170, 92.785641)    (171, 92.988788)    (172, 92.953346)    (173, 92.899345)    (174, 92.951445)    (175, 93.014838)    (176, 93.026018)    (177, 92.981297)    (178, 93.018527)    (179, 92.975707)    (180, 93.016738)    (181, 93.024117)    (182, 93.024117)    (183, 93.009248)    (184, 92.984986)    (185, 93.085609)    (186, 92.975707)    (187, 93.052068)    (188, 93.095001)    (189, 93.027919)    (190, 93.083820)    (191, 93.087510)    (192, 92.975707)    (193, 93.063249)    (194, 93.020428)    (195, 92.930985)    (196, 93.135921)    (197, 93.100591)    (198, 93.111771)    (199, 93.115461)    (200, 93.130331)    (201, 93.206804)    (202, 93.121051)    (203, 93.122952)    (204, 93.063249)    (205, 93.171362)    (206, 93.271986)    (207, 93.076330)    (208, 93.156493)    (209, 93.204903)    (210, 93.225363)    (211, 93.197413)    (212, 93.162083)    (213, 93.221674)    (214, 93.242134)    (215, 93.141511)    (216, 93.227264)    (217, 93.201214)    (218, 93.247724)    (219, 93.271986)    (220, 93.234755)    (221, 93.245935)    (222, 93.208593)    (223, 93.307427)    (224, 93.262706)    (225, 93.271986)    (226, 93.298036)    (227, 93.275675)    (228, 93.260805)    (229, 93.130331)    (230, 93.184443)    (231, 93.251525)    (232, 93.214183)    (233, 93.296247)    (234, 93.283166)    (235, 93.296247)    (236, 93.251525)    (237, 93.318608)    (238, 93.324198)    (239, 93.314806)    (240, 93.385690)    (241, 93.305527)    (242, 93.333477)    (243, 93.335378)    (244, 93.255215)    (245, 93.337167)    (246, 93.329788)    (247, 93.286855)    (248, 93.372609)    (249, 93.331577)    (250, 93.314806)
  };
\addplot[]
  coordinates {
(0, 24.432432)  (1, 44.549688)(2, 56.596855)(3, 62.102859)(4, 66.083265)(5, 68.563395)(6, 71.450773)(7, 70.905093)(8, 73.653296)(9, 74.766428)(10, 75.325960)(11, 76.262903)(12, 75.887467)(13, 77.394473)(14, 78.915509)(15, 78.533935)(16, 77.963218)(17, 78.724613)(18, 79.106624)(19, 79.873274)(20, 79.106624)(21, 79.297958)(22, 80.065484)(23, 80.643426)(24, 80.643426)(25, 79.681283)(26, 79.681283)(27, 79.489511)(28, 80.450560)(29, 81.029814)(30, 80.836510)(31, 81.611037)(32, 81.805215)(33, 81.417077)(34, 81.611037)(35, 81.611037)(36, 81.417077)(37, 82.779383)(38, 82.194226)(39, 82.194226)(40, 81.999611)(41, 82.584112)(42, 83.366506)(43, 82.779383)(44, 82.584112)(45, 83.955593)(46, 83.759013)(47, 83.759013)(48, 83.955593)(49, 84.744100)(50, 84.744100)(51, 83.759013)(52, 84.546646)(53, 83.759013)(54, 84.546646)(55, 84.941772)(56, 83.955593)(57, 83.759013)(58, 82.974872)(59, 84.152393)(60, 84.349410)(61, 84.744100)(62, 84.744100)(63, 85.139662)(64, 85.139662)(65, 85.734641)(66, 84.546646)(67, 85.933403)(68, 85.536096)(69, 85.139662)(70, 86.132384)(71, 85.734641)(72, 85.734641)(73, 86.730633)(74, 85.734641)(75, 85.139662)(76, 85.536096)(77, 85.734641)(78, 85.933403)(79, 86.331582)(80, 85.734641)(81, 85.139662)(82, 86.730633)(83, 85.139662)(84, 85.139662)(85, 86.930485)(86, 86.530999)(87, 86.530999)(88, 85.139662)(89, 86.132384)(90, 86.530999)(91, 86.530999)(92, 86.132384)(93, 86.530999)(94, 86.331582)(95, 86.530999)(96, 86.730633)(97, 86.132384)(98, 86.530999)(99, 87.130555)(100, 86.930485)(101, 86.331582)(102, 86.730633)(103, 86.530999)(104, 86.530999)(105, 85.734641)(106, 86.730633)(107, 86.530999)(108, 86.730633)(109, 85.337770)(110, 86.331582)(111, 85.734641)(112, 86.132384)(113, 86.530999)(114, 86.331582)(115, 86.331582)(116, 86.132384)(117, 86.331582)(118, 86.530999)(119, 86.930485)(120, 86.930485)(121, 85.933403)(122, 86.331582)(123, 86.930485)(124, 86.930485)(125, 86.530999)(126, 87.330843)(127, 87.130555)(128, 85.933403)(129, 86.132384)(130, 86.331582)(131, 85.933403)(132, 87.130555)(133, 86.530999)(134, 86.730633)(135, 86.331582)(136, 85.734641)(137, 86.730633)(138, 86.930485)(139, 86.930485)(140, 86.530999)(141, 87.531349)(142, 87.330843)(143, 86.331582)(144, 87.130555)(145, 86.530999)(146, 86.930485)(147, 87.531349)(148, 87.130555)(149, 86.530999)(150, 86.930485)(151, 87.933014)(152, 87.330843)(153, 87.330843)(154, 86.930485)(155, 86.930485)(156, 87.330843)(157, 87.130555)(158, 87.732072)(159, 87.330843)(160, 86.730633)(161, 87.531349)(162, 87.732072)(163, 86.930485)(164, 87.130555)(165, 87.933014)(166, 87.531349)(167, 87.330843)(168, 87.130555)(169, 87.933014)(170, 87.933014)(171, 87.330843)(172, 87.732072)(173, 87.531349)(174, 87.531349)(175, 87.531349)(176, 87.732072)(177, 87.732072)(178, 87.732072)(179, 87.732072)(180, 88.134173)(181, 87.531349)(182, 87.732072)(183, 87.933014)(184, 87.531349)(185, 87.531349)(186, 86.530999)(187, 88.537144)(188, 87.330843)(189, 86.930485)(190, 88.335550)(191, 87.330843)(192, 87.933014)(193, 87.130555)(194, 87.130555)(195, 87.330843)(196, 87.732072)(197, 88.134173)(198, 87.330843)(199, 88.134173)(200, 87.933014)(201, 87.130555)(202, 87.531349)(203, 86.930485)(204, 87.130555)(205, 87.330843)(206, 87.732072)(207, 87.933014)(208, 87.130555)(209, 87.933014)(210, 87.330843)(211, 87.933014)(212, 87.933014)(213, 87.732072)(214, 88.335550)(215, 87.531349)(216, 87.732072)(217, 87.732072)(218, 87.933014)(219, 88.134173)(220, 86.930485)(221, 87.933014)(222, 87.732072)(223, 87.732072)(224, 88.134173)(225, 87.933014)(226, 88.335550)(227, 88.134173)(228, 87.933014)(229, 88.537144)(230, 87.933014)(231, 87.531349)(232, 87.933014)(233, 87.933014)(234, 87.330843)(235, 87.330843)(236, 87.531349)(237, 87.933014)(238, 86.730633)(239, 87.933014)(240, 87.330843)(241, 87.933014)(242, 88.134173)(243, 87.933014)(244, 88.537144)(245, 87.933014)(246, 87.732072)(247, 87.933014)(248, 87.531349)(249, 88.537144)(250, 88.335550)
  };

@@

The training performance is intentionally dashed, as  what interests us is the model's generalization power. It quickly learns in the first 50 epochs and reaches a plateau around the 100 mark. It achieves 93.3\% performance on the training set and 88.3\% on unseen data.
